#Peer Assignment 1 of Course on Reproducible Research

## Data Description
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

date: The date on which the measurement was taken in YYYY-MM-DD format

interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset

## Step 1: Loading and preprocessing the data
``` {r}
activity<-read.csv ("activity.csv")
summary(activity)
validActivityData<-activity[!is.na(activity$steps), ]
summary(validActivityData)
```

## Step 2: Calculating mean total number of steps taken per day
### 2.1 Calculating total steps per day
```{r }
library(data.table)
activityTable<-data.table(validActivityData)
activityTable[, sum(steps), by=date]
```
### 2.2 Plotting a histogram of total steps per day
```{r}
library (ggplot2)
data<-aggregate (steps~date, validActivityData, sum)
qplot (data$steps)
```
<p>
### 2.3 Mean and median of steps per day
``` {r}
activityTable[, mean(steps), by=date]
activityTable[, median(steps), by=date]
```

## Step 3 Finding the Average Daily Pattern
### Step 3.1 Time-Series Plot
```{r}
dt<-activityTable[, mean(steps), by=interval]
qplot (dt$interval, dt$V1, type="l")
max (dt$V1)
dt[dt$V1>206,]
```
<p>
From the plot, it appears that maximum activity occurs in the 5-minute interval starting at 8:35 am

## Step 4 Imputing Missing Values
### 4.1 Number of missing values in the original data
Number of rows with NA in original data is `r nrow(activity[is.na(activity$steps),])`

### 4.2 Imputing missing values

We will create a new dataset with missing values filled in using the mean values for the corresponding time interval

First, we find the means for specific intervals across all days
```{r}
meansForIntervals<-aggregate (steps~interval, validActivityData, mean)
summary(meansForIntervals)
```

Now, we create the new dataset by imputing values
```{r}
changedData=merge (activity, meansForIntervals, by="interval")
for (i in 1:nrow(changedData)) {if (is.na(changedData$steps.x[i])) changedData$steps.x[i] = changedData$steps.y[i]}
summary(changedData)
```

### 4.3 Calculating mean, median and plotting histogram of the changed data

### 4.3.1 Plotting a histogram of total steps per day
```{r}
library (ggplot2)
data<-aggregate (steps.x~date, changedData, sum)
qplot (data$steps.x)
```
<p>
### 4.3.2 Mean and median of steps per day
``` {r}
library(data.table)
activityTable<-data.table(changedData)
activityTable[, mean(steps.x), by=date]
activityTable[, median(steps.x), by=date]
```
